### 2021-02-03 static/tree.tex
writing draft

\documentclass{article}
\usepackage{tikz}
\usetikzlibrary{trees}
\tikzstyle{densely dashdotted}=      [dash pattern=on 3pt off 1pt on \the\pgflinewidth off 1pt]
\tikzset{
  treenode/.style = {shape=circle, inner sep=1pt,
draw, align=center, fill, black
%top color=white, bottom color=blue!20
  },
  payoff/.style    = {align=center, inner sep=0.1em, text width=1.5em},
  left side node/.style={above left, inner sep=0.1em},
  right side node/.style={above right, inner sep=0.1em}
}

\begin{document}
\begin{tikzpicture}
  [
%grow                    = right,
sibling distance        = 25em,
level 2/.style          = {sibling distance=2cm},
level 3/.style          = {sibling distance=3.5cm},
level 4/.style          = {sibling distance=2cm},
level distance          = 6em,
every node/.style       = {font=\footnotesize},
%sloped
  ]
  \node [treenode] {} child{node[payoff] {$S_{L}, S_{T}$}
  child{node [payoff] {\\ $S_{L}, S_{M} - S_{T}, S_{M}$} edge from parent[densely dashdotted] node[left side node, rotate=35] {$\frac{k-1}{N-2}\frac{k-2}{N-3}$}}
  child{node [payoff] {\\ $S_{L}, S_{R} - S_{T}, S_{M}$} edge from parent[densely dashdotted] node[left side node, rotate=45] {$\frac{k-1}{N-2}\frac{N-k-1}{N-3}$}}
  child{ node [payoff] {\\ $S_{L}, S_{M} - S_{T}, S_{R}$} edge from parent[densely dashdotted] node[right side node, rotate=-45] {$\frac{N-k-1}{N-2}\frac{k-1}{N-3}$}}
  child{ node [payoff] {\\ $S_{L}, S_{R} - S_{T}, S_{R}$} edge from parent[densely dashdotted] node[right side node, rotate=-35, xshift=-.5cm] {$\frac{N-k-1}{N-2}\frac{N-k-2}{N-3}$}}
  edge from parent node[left side node] {$\frac{1}{1 - N}$}}
  child{node [treenode] {} 
  child{node [payoff] {\\ $S_{L}, S_{M} - S_{T}, S_{M}$}
  child{node [payoff] {\\ $S_{L}, S_{T}$} edge from parent[densely dashdotted] node[left side node] {$\frac{1}{N-3}$}}
  child{node [treenode] {} 
  child{node [payoff] {\\ $S_{L}, S_{M} -  S_{T}, S_{M}$} edge from parent[densely dashdotted] node[left side node, rotate=35] {$\frac{k-2}{N-3}\frac{k-3}{N-4}$}}
  child{node [payoff] {\\ $S_{L}, S_{R} -  S_{T}, S_{M}$} edge from parent[densely dashdotted] node[left side node, rotate=45] {$\frac{N-k-1}{N-3}\frac{k-2}{N-4}$}}
  child{ node [payoff] {\\ $S_{L}, S_{M} - S_{T}, S_{R}$} edge from parent[densely dashdotted] node[right side node, rotate=-45] {$\frac{k-2}{N-3}\frac{N-k-1}{N-4}$}}
  child{ node [payoff] {\\ $S_{L}, S_{R} - S_{T}, S_{R}$} edge from parent[densely dashdotted] node[right side node, rotate=-35, xshift=-.5cm] {$\frac{N-k-1}{N-3}\frac{N-k-2}{N-4}$}}
  edge from parent node[right side node] {$1 - \frac{1}{N-3}$}}
  edge from parent node[left side node, rotate=35] {$\frac{k-1}{N-2}\frac{k-2}{N-3}$}}
  child{node [payoff] {\\ $S_{L}, S_{R} -  S_{T}, S_{M}$} edge from parent[densely dashdotted] node[left side node, rotate=45] {$\frac{k-1}{N-2}\frac{N-k-1}{N-3}$}}
  child{ node [payoff] {\\ $S_{L}, S_{M} - S_{T}, S_{R}$} edge from parent[densely dashdotted] node[right side node, rotate=-45] {$\frac{N-k-1}{N-2}\frac{k-1}{N-3}$}}
  child{ node [payoff] {\\ $S_{L}, S_{R} - S_{T}, S_{R}$} edge from parent[densely dashdotted] node[right side node, rotate=-35, xshift=-.5cm] {$\frac{N-k-1}{N-2}\frac{N-k-2}{N-3}$}
  edge from parent node[right side node, xshift=-3cm, yshift=2.1cm] {$1 - \frac{1}{1 - N}$}}};
\end{tikzpicture}
\end{document}