### 2021-05-25 main.tex
first draft to CH

\documentclass[11pt]{article}
\usepackage{times}
\usepackage{amsmath,amsthm,amssymb,setspace,enumitem,epsfig,titlesec,verbatim,color,array,eurosym,multirow}
\usepackage[sort&compress]{natbib}
\usepackage[footnotesize,bf]{caption}
\usepackage[margin=2.5cm, includefoot, footskip=30pt]{geometry}
\usepackage{standalone}
\usepackage{tikz}
\usepackage{subcaption}
\usepackage{hyperref}
\usepackage{tabularx}
\usepackage{booktabs}
\usepackage[ruled,vlined]{algorithm2e}
\smallskip % Erlaubt kleine Abstaende zwischen Paragraphen, falls es dem Seitenlayout hilft
\renewcommand{\baselinestretch}{1.3}
\newcommand{\R}{\mathbb{R}}

\definecolor{darkblue}{rgb}{0,0,.8}
\definecolor{darkgreen}{rgb}{0,0.5,0.1}
\newcommand{\matlabfunc}[1]{\textcolor{darkblue}{#1}}
\newcommand{\matlabcomment}[1]{\textcolor{darkgreen}{#1}}
\newcommand{\christian}[1]{\textcolor{blue}{\textbf{CH}: #1}}
\newcommand{\alex}[1]{\textcolor{red}{\textbf{AL}: #1}}

%% Adding shortcut commands to refer to our figures %%
\newcommand{\FigEvoProc}{{\bf Fig.~1}}
\newcommand{\FigInvAnalysis}{{\bf Fig.~2}}
\newcommand{\FigResultsOverPara}{{\bf Fig.~3}}

\titleformat{\section}{\sffamily \fontsize{12}{14}\bfseries}{\thesection}{1em}{}
\titleformat{\subsection}{\sffamily \fontsize{11.5}{11.5}\bfseries}{\thesubsection}{1em}{}

\usepackage{tikz}
\usetikzlibrary{arrows}

\tikzset{
  treenode/.style = {align=center, inner sep=0pt, text centered,
    font=\sffamily},
  arn_n/.style = {treenode, circle, white, font=\sffamily\bfseries, draw=black, inner sep=-6pt,
    fill=black, text width=1.5em},% arbre rouge noir, noeud noir
  arn_r/.style = {treenode, circle, red, 
    text width=1.5em, very thick, inner sep=4pt},% arbre rouge noir, noeud rouge
  arn_x/.style = {treenode, rectangle, draw=black,
    minimum width=0.5em, minimum height=0.5em}% arbre rouge noir, nil
}

\newtheoremstyle{plainCl1}% name
{9pt}%      Space above, empty = 'usual value'
{9pt}%      Space below
{\it}% 	   Body font
{}%         Indent amount (empty = no indent, \parindent = para indent)
{\bfseries}% Thm head font
{.}%        Punctuation after thm head
{0.2cm}% Space after thm head: \newline = linebreak
{}%         Thm head spec

\newtheoremstyle{plainCl2}% name
{9pt}%      Space above, empty = 'usual value'
{9pt}%      Space below
{\it}% 	   Body font
{}%         Indent amount (empty = no indent, \parindent = para indent)
{\bfseries}% Thm head font
{$'$.}%        Punctuation after thm head
{0.2cm}% Space after thm head: \newline = linebreak
{}%         Thm head spec

\newcommand{\splitatcommas}[1]{%
  \begingroup
  \begingroup\lccode`~=`, \lowercase{\endgroup
    \edef~{\mathchar\the\mathcode`, \penalty0 \noexpand\hspace{0pt plus 1em}}%
  }\mathcode`,="8000 #1%
  \endgroup
}

\theoremstyle{plainCl1}
\newtheorem{Claim}{Claim}
\newtheorem{Thm}{Theorem}
\newtheorem{Prop}{Proposition}
\newtheorem*{Lem}{Lemma}
\newtheorem{Cor}{Corollary}
\newtheorem*{Def}{Definition}

\theoremstyle{plainCl2}
\newtheorem{Claim2}{Claim}

\title{
\bf  \sffamily \LARGE Evolution of cooperation among individuals with limited payoff memory\\}
\date{}
\author{Christian Hilbe, Nikoleta E. Glynatsi, Alex McAvoy}

\begin{document}
\maketitle

\begin{abstract}
Repeated games are vastly used in evolutionary game theory to explain
cooperation in a variety of environments. Although existing models have shaped
our understanding of human cooperation, the often work with idealized
assumptions. In an evolutionary process, individuals imitate other individuals
of the population based on their fitness. It is commonly assumed that an
individual computes their fitness after interacting with a representative sample
of the population, and remembering all the interactions they participated in. In
real life, we do not always remember all our interactions, instead we rather
recall our most recent ones.

Here, we introduce a framework that allows individuals to estimate their fitness
based on a minimum of social information. We explore the difference between the
classical framework and ours using computer simulations. We present results for
the most commonly used classes of symmetric \(2 \times 2\) games. The
simulations show that, in the prisoner's dilemma individuals with limited memory
tend to adopt less generous strategies and they achieve less cooperation than in
the classical scenario. In contrast, in the stag-hunt and the harmony game, the
impact of memory is less striking. Here individuals with limited memory perform
nearly as well as individuals with full memory. Finally, we observe that the
snowdrift game is the only game for which cooperation can be underestimated in
the classical scenario.
\end{abstract}

\section{Introduction}

One of the most important applications of evolutionary game theory is the
evolution of cooperation. Why is it that some individuals choose to help others,
increasing their payoff, at the expense of decreasing one's own? In evolutionary
game theory individuals are not required to be rational, instead they adapt
strategies based on mutation and exploration. Strategies are more likely to
spread if they have a high fitness either because the individuals who adopt them
have more offspring, or because they are imitated more often~\cite{Wu2015}. The
fitness of a strategy is not constant but depends on the composition of the
population. Individuals interact based on their strategies with other members of
the population and the payoffs they yield are translated into fitness.

It is commonly assumed that individuals compute their fitness after interacting
with a representative sample of the population, and remembering all the
interactions they participated in~\cite{Nowak2006, nowak2004emergence,
Hilbe2013, Hilbe2018,Reiter2018, garcia2018no, Nowak1992tit, Baek2016}. This
implies that individuals have a perfect memory. However, when modelling how
individuals make decisions in each turn they are assumed to have very limited
memory. To be precise, most of the studies in the literature focus on naive
subjects who can only choose from a restricted set of
strategies~\cite{Nowak1992tit}, or who do not remember anything beyond the
outcome of the very last round~\cite{Baek2016}. Note that there are a few
notable exceptions~\cite{Hauert1997, Stewart2016}.

Studies that make use of this classical framework report that cooperation can
substantially evolve. However, the inconsistency in the memory size of players
leads us to question the robustness of our understanding of cooperation. To this end, we
propose a framework in which individuals, similar to the decisions at each turn,
estimate their fitness based on a minimum of information. Though we are not the
first to question the assumptions of estimating fitness~\cite{Roca2006}, we are
the first to explore the effect of memory.

We first consider two extreme scenarios, the classical scenario
and the alternative scenario where individuals update their strategies
only based on the very last payoff they obtained. We observe that individuals 
with limited memory tend to adopt less generous strategies and they achieve less
cooperation when interacting in a prisoner's dilemma. We obtain similar results when we
consider that individuals update their strategies based on more
information. More specifically, up to the last two payoffs they obtained when
interacting with up to two different members of the population. We extend our
approach to the rest of the symmetric \(2 \times 2\) games.

The remainder of the paper is organized as follows. In
section~\ref{section:model} we describe the model. In
section~\ref{section:results} we present the results of the simulations, and
in section~\ref{section:conclusions} we outline the main conclusions.

\section{Model Setup}\label{section:model}

We consider a population of \(N\) players~\footnote{The terms ``player'' and
``individual'' are used interchangeably here.} where \(N\) is even and mutations
are sufficiently rare. Therefore, at any point in time there are at most two
different strategies present in the population; a \textit{resident} strategy and
a \textit{mutant} strategy. To describe how strategies spread we use a pairwise
comparison process~\cite{Traulsen2006}.
Each step of the evolutionary process consists of two
stages, a game stage and an updating stage.

In the game stage each individual is randomly matched with some other individual
in the population. They engage in a match where each subsequent turn
occurs with a fixed probability $\delta$. At each turn players choose independently to
either cooperate (\(C\)) or to defect (\(D\)), and the payoffs of the turn
depend on both their decisions. If both players cooperate they receive the
reward payoff \(R\), whereas if both defect they receive the punishment payoff
\(P\). If one cooperates but the other defects, the defector receives the
temptation payoff \(T\), whereas the cooperator receives the sucker's
payoff \(S\). We denote the feasible payoff of each turn as \(\mathcal{U} =
\{R, S, T, P\}\).
We assume that individuals use \textit{reactive strategies} to make
decisions in each turn. Reactive strategies are a set of memory-one strategies
that only take into account the previous action of the opponent. They can be
written explicitly as a vector in \(\R_{3}\), more specifically, a reactive
strategy \(s\) is given by \(s=(y, p, q)\). The parameter \(y\) is the probability that
the strategy opens with a cooperation and \(p, q\) are the probabilities that
the strategy cooperates given that the opponent cooperated and defected
equivalently.

In the updating stage, two players are randomly drawn from the population, a
`learner' and a `role model'. Given the learner's payoff $u_L\!\in\!
\mathcal{U}$ and the role model's payoff $u_{RL}\!\in\! \mathcal{U}$,
the learner adopts the role model's strategy with probability,

\begin{equation} \label{Eq:rho}
  \rho(u_{L}, u_{RM}) = \frac{1}{1\!+\! \exp^{\!-\!\beta (u_{RM}\!-\!u_{L})}}.
\end{equation}

where $\beta\!\ge\!0$ is the strength of selection. For small values of $\beta$
the imitation probability is independent of the strategies of the involved
players. As the value of $\beta$ increases, the more likely it is that the
learner adopts only strategies that yield a higher payoff.
Conventionally the updating payoffs of the learner and the role model are based
on their expected payoffs. A player's expected payoff is the mean payoff the
player yields after engaging in matches of multiple turns with each member of
the population. At each match a player bases their next turn decision only on
the previous action of the opponent, however, the same player bases their
expected payoffs on the outcomes of all their matches. Thus, a player is assumed
to have limited and perfect memory at the same time. We propose a new a set of
updating payoffs where it is also assumed that a player has limited memory.
We referee to these as the limited memory payoffs.

The evolutionary step is repeated until either the mutant strategy goes
extinct, or until it fixes in the population. If the mutant fixes in the
population then the mutant strategy becomes the new resident strategy. After
either outcome we introduce a new mutant strategy uniformly chosen from all
reactive strategies at random, and we set the number of mutants to $1$. This
process of mutation and fixation/extinction is then iterated many times.

In order to account for the effect of the updating payoffs we simulate the
evolutionary process and record which strategies the players adopt over time
based on (i) the expected payoffs (ii) the limited memory payoffs. We compare
the difference in the cooperation rate within the resident population for the
two approaches. To account for the various types of social behaviour we also
present results on multiple social dilemmas.

\section{Results}\label{section:results}
\subsection{Updating payoffs based on the last round with another member of the population}\label{section:donation}

In this section we explore the case where the updating payoffs are based on the
last round payoff achieved against another member of the population and we compare
this to the classical scenario of the expected payoffs. We assume that each pair
of players interacts in a donation game. The donation game is a special case of
the prisoner's dilemma. Each player can choose to cooperate by providing a
benefit \(b\) to the other player at their cost \(c\), with \(0 < c < b\).
Thus, the feasible payoffs in each round are \(\mathcal{U} = \{b-c, -c, b, 0\}\).

Figure~\ref{fig:expected_and_stochastic_for_donation} shows simulations results
for the described process of section~\ref{section:model}.
Figure~\ref{fig:expected_and_stochastic_for_donation} depicts the evolving
conditional cooperation probabilities $p$ and $q$. The discount factor~$\delta$
is comparably high, thus we do not report the opening move \(y\) as it
is a transient effect. The left panels correspond to the
standard scenario considered in the literature, it considers players who use
expected payoffs to update their strategies. The right panel shows the scenario
considered herein, in which players update their strategies based on their last
round’s payoff. The top panels assume a benefit \(b\) of 3 whereas the bottom
assume a benefit of 10.

The figure suggests that when updating is based on expected payoffs players
tend to be more generous and more cooperative. The $q$-values of the resident
strategies are on average higher in the case of the expected payoffs. The
players will occasionally forgive a defection more often if their fitness
depends on interacting with every member of the population. On the other hand,
when social interactions are limited they are less forgiving. The average
cooperation rate for each simulation is calculated as the average
cooperation rate within the resident population. In the case of the expected
payoffs, regardless the value of benefit, the average cooperation rate is
strictly higher than that of the last round payoffs. The difference based on the
two methods is statistically significant, and in the case of $b=10$ the
average cooperation of resident strategies drops from 97\% to 57\%.

\begin{figure}[!htbp]
    \centering
    \includegraphics[width=.70\textwidth]{static/expected_and_stochastic_for_donation_game.pdf}\vspace{-3cm}
    \includegraphics[width=.70\textwidth]{static/expected_and_stochastic_for_donation_game_b_10.pdf}
    \caption{{\bf Evolutionary dynamics under expected payoffs and last round with one interaction payoffs.} 
    We have run two simulations of the evolutionary process described in
    section~\ref{section:model} for $t\!=\!10^7$ time steps. For each time step,
    we have recorded the current resident population ($y,p,q$). Since simulations
    are run for a relatively high continuation probability of $\delta\!=\!0.999$, we
    do not report the players' initial cooperation probability $y$. The graphs show
    how often the resident population chooses each combination ($p,q$) of
    conditional cooperation probabilities in the subsequent rounds. ({\bf A}) If
    players update based on their expected payoffs, the resident population
    typically applies a strategy for which $p\!\approx\!1$ and
    $q\!\le\!1\!-\!c/b\!=\!0.9$. ({\bf B})
    When players update their strategies based on their realized payoffs in the last
    round, there are two different predominant behaviors. The resident population
    either consists of defectors (with $p\!\approx\!q\!\approx\!0$) or of
    conditional cooperators. In the latter case, the maximum level of $q$ consistent
    with stable cooperation is somewhat smaller compared to the expected-payoff
    setting, $q\!<\!0.5$. The cooperation rate within the resident population
    (averaged over all games and over all time steps) is close to 100\%. Parameters:
    $N\!=\!100$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.999$.}
    \label{fig:expected_and_stochastic_for_donation}
\end{figure}

We further explore the effect of benefit in
Figure~\ref{fig:cooperation_rate_over_benefit}. The figure suggests that
expected payoffs always yield a higher cooperation rate. In the case of expected
payoffs we observe that the cooperation rate increases as the value of the
benefit gets higher. In comparison for the limited memory payoffs, the
cooperation rate remains unchanged at approximately 50\% once \(b=5\).

\begin{figure}[!htbp]
  \centering
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=\textwidth]{static/expected_for_beta.pdf}
    \includegraphics[width=\textwidth]{static/stochastic_for_beta.pdf}
  \end{subfigure}%
  \begin{subfigure}{.5\textwidth}
    \centering
    \includegraphics[width=\textwidth]{static/cooperation_rate_over_b.pdf}
  \end{subfigure}
  \caption{{\bf The evolution of cooperation for different benefit values.} 
  We vary the benefit of defection $b$. In all cases, expected payoffs appear to
  overestimate the average cooperation rate the population achieves. ({\bf A})
  the probabilities \(p, q\) for resident population over \(10^7\) time steps
  for each benefit value. ({\bf B}) The cooperation rate within the resident population
  (averaged over all games and over all time steps) over the benefit.
  Unless
  explicitly varied, the parameters of the simulation are $N\!=\!100$,
  $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run
  for $t\!=\!5\times 10^7$ time steps for each parameter
  combination.}\label{fig:cooperation_rate_over_benefit}
\end{figure}

We also investigate the effect of the strength of selection $\beta$.
Figure~\ref{fig:cooperation_rate_over_betas} illustrates results for various
runs of the evolutionary process. For weak selection, \(\beta < 1\), we observe
that the two methods yield similar results, however, as \(\beta\) increases
there is variation in the evolving populations. In the case of expected payoffs
the resident populations become more cooperative as \(\beta\)
increases, whereas in the case of limited memory payoffs, the resident
populations become more defective.

\begin{figure}[!htbp]
  \centering
  \begin{subfigure}{.65\textwidth}
    \centering
    \includegraphics[width=\textwidth]{static/expected_for_selection_strenght.pdf}
    \includegraphics[width=\textwidth]{static/stochastic_for_selection_strenght.pdf}
  \end{subfigure}%
  \begin{subfigure}{.35\textwidth}
    \vspace{.3cm}
    \includegraphics[width=\textwidth]{static/cooperation_rate_over_betas.pdf}
  \end{subfigure}
\caption{{\bf The evolution of cooperation for different selection strength values.}
We vary the selection strength $\beta$. In all cases, stochastic payoff
evaluation tends to reduce the evolving cooperation rates. ({\bf A}) the
probabilities \(p, q\) for resident population over \(10^7\) time steps for each
\(\beta\) value. ({\bf B}) The cooperation rate within the resident population
(averaged over all games and over all time steps) over \(\beta\).
Unless explicitly varied, the parameters of the simulation are $N\!=\!100$,
$b\!=\!3$, $c\!=\!1$, $\beta\!=\!1$, $\delta\!=\!0.99$. Simulations are run for
$t\!=\!5\times 10^7$ time steps for each parameter
combination.}\label{fig:cooperation_rate_over_betas}
\end{figure}

\subsection{Effect of updating payoffs in different social dilemmas}\label{section:2_by_2_games}

In the previous section we gained insights into the effects of the updating
payoffs, and into how parameters such as the benefit and the strength of
selection can intensify them. We investigated these effects by using the
donation game. In order to broaden our understanding of the updating payoffs on
different forms of possible human interactions we extend our approach to all
\(2 \times 2\) symmetric games. More specifically, we apply our analysis to four
different classes of games; the harmony, the snowdrift, the prisoner's dilemma
and the stag-hunt games.

We compare the results of the evolutionary process when the updating payoffs are based on the expected
payoffs, and on the last round payoffs for all the four possible classes of
games (Figures~\ref{fig:expected_two_by_two}
and~\ref{fig:last_round_two_by_two}). In Figure~\ref{fig:expected_two_by_two}
each sub figure represents a run of evolutionary process for a different set of
values for \(\mathcal{U}\). Without loss of generality we set \(R=1\) and
\(P=0\)~\cite{Martinez2012, Roca2009}, and we vary the values of the temptation
payoff \(T\) (across the \(x\)-axis) and of the sucker's payoff \(S\) (across
the \(y\)-axis). Starting at the upper left corner and proceeding clockwise the
quadrants correspond to the harmony, the snowdrift, the prisoner's dilemma and
the stag-hunt games.

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\textwidth]{static/expected_two_by_two_games.pdf}
  \caption{{\bf Evolutionary dynamics under expected payoffs for various social dilemmas.}
  We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across
  the \(x\) axis, and  \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\) across the
  \(y\) axis. ({\bf A}) The top left quadrant where \(S > 0\) and \(T \leq 1\)
  corresponds to the harmony game. The preference ordering for the harmony game
  is \(R > T > S > P\). ({\bf B}) The top right quadrant where \(S > 0\) and \(T > 1\)
  corresponds to the snowdrift game. The preference ordering for the snowdrift game
  is \(T > R > S > P\). ({\bf C}) The bottom left quadrant where \(S < 0\) and \(T \leq 1\)
  corresponds to the stag-hunt game. The preference ordering for the stag-hunt game
  is \(R > T > P > S\). ({\bf D}) The bottom right quadrant where \(S < 0\) and \(T > 1\)
  corresponds to the prisoner's dilemma game. The preference ordering for the prisoner's dilemma game
  is \(T > R > P > S\).
  Unless explicitly varied, the parameters of the simulation
  are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}
  \label{fig:expected_two_by_two}
\end{figure}

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\textwidth]{static/stochastic_two_by_two_games.pdf}
  \caption{{\bf Evolutionary dynamics under last round payoffs for various social dilemmas.} 
  We vary the temptation payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\) across
  the \(x\) axis, and  \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\) across the
  \(y\) axis. ({\bf A}) The top left quadrant where \(S > 0\) and \(T \leq 1\)
  corresponds to the harmony game. The preference ordering for the harmony game
  is \(R > T > S > P\). ({\bf B}) The top right quadrant where \(S > 0\) and \(T > 1\)
  corresponds to the snowdrift game. The preference ordering for the snowdrift game
  is \(T > R > S > P\). ({\bf C}) The bottom left quadrant where \(S < 0\) and \(T \leq 1\)
  corresponds to the stag-hunt game. The preference ordering for the stag-hunt game
  is \(R > T > P > S\). ({\bf D}) The bottom right quadrant where \(S < 0\) and \(T > 1\)
  corresponds to the prisoner's dilemma game. The preference ordering for the prisoner's dilemma game
  is \(T > R > P > S\).
  Unless explicitly varied, the parameters of the simulation
  are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}
  \label{fig:last_round_two_by_two}
\end{figure}

The harmony game describes an interaction without conflict. In the harmony game
it is in the best interest for both players to cooperate, and in an evolutionary
process cooperators prevail. We observe that for most harmony game runs the
resident population overwhelmingly applies a strategy for which \(p\) and \(q\)
are \(\approx 1\) (Figure~\ref{fig:expected_two_by_two}). The snowdrift game
describes a situation similar to that of the prisoner's dilemma; cooperation
results in a benefit to the opposing player but entailing a cost to the
cooperator. However, in the snowdrift game individuals obtain immediate direct
benefits from the cooperative acts which leads to \(S > P\). In the snowdrift
game more cooperation emerges compared to the prisoner's dilemma. Defection is
never a resident strategy and for the same values of temptation the overall
\(q-\) values are higher. The last class of games we present are for the stag-hunt
game. In the stag-hunt game there is an equilibrium in which both players
cooperate as well as one in which both defect. For small values of the
temptation and the sucker's payoffs we observe that cooperation prevails similar
to the harmony game. However, in the case where the temptation and the reward
payoffs are equal the resident population either consists of defectors or
cooperators.

Although the patterns of the evolved populations remain similar if we consider
that players update their strategies based on their last round payoff
(Figure~\ref{fig:last_round_two_by_two}), there are still dissimilarities
between the two approaches. To better understand these dissimilarities, we
explore the cooperation rates for each of the cases of the social dilemmas we
have presented (Figure~\ref{fig:cooperation_rates_two_by_two_part_one}).

In the harmony game the cooperation rates for both approaches are similar. The
biggest difference occurs for \(T=1\), nevertheless, it is less than 10\%. There
are two instances for which the cooperation rates are higher for the limited
memory payoffs compared to the expected. These are for \(T=2.2, S=0.4\) and for
\(T=2.2, S=0.8\). It is only in the snowdrift game that cooperation is
overestimated when a minimum of information is considered. Players are slightly
more likely to cooperate after a cooperation and to forgive after being at the
receiving end of a defection. For the rest of the cases the classical scenario
overestimates the average cooperation rate, supporting the results of the
previous section. There is a substantial difference in the cooperation rates
for the prisoner's dilemma class. We observe that for \(S<-0.4\) cooperation
almost never emerges if players imitate based only on the last round payoff.

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=.4\textwidth]{static/expected_two_by_two_games_cooperation.pdf}
  \includegraphics[width=.4\textwidth]{static/stochastic_two_by_two_games_cooperation.pdf}
  \caption{\textbf{Cooperation rates for various social dilemmas.} ({\bf A}) If
  players update based on their expected payoffs. ({\bf B}) If
  players update based on their last round payoffs. We vary the temptation
  payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\)
  across the \(x\) axis, and \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\)
  across the \(y\) axis. Unless explicitly varied, the parameters of the simulation
  are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$.}\label{fig:cooperation_rates_two_by_two_part_one}
\end{figure}

So far we have explored the difference between the expected payoffs and the last
round payoffs. In order to explore further the effect of limited memory we allow
individuals to remember more. More precisely, up to two interactions and up to
the last two rounds. In total we present results for three more updating
payoffs. These are the payoffs when individuals consider the last two rounds
with another member of the population, the last round with two members of the
population, and the last two rounds with two members of the population.
Similar to the last round payoff, we use simulations and record which strategies
the players adopt over time based
(Appendix~\ref{appendix:further_simulation_results}) and we compare the
evolving cooperation rates to those of the expected payoffs
(Figure~\ref{fig:cooperation_rates_two_by_two_part_two}).

In the case of the last two rounds payoff we make two observations. Initially,
for the harmony, the stag-hunt and the prisoner's dilemma classes the average
cooperation rates are higher than the previous case of the last round payoff.
The rates remain strictly less than in the expected payoffs, however, in the
case of the prisoner's dilemma we observe a significant increase. Secondly, for
the snowdrift class there are more instances for which the cooperation rates are
higher than in the expected payoffs. This could indicate that in the case of the
snowdrift game the expected payoffs underestimate cooperation.
Once players interact with two other members of the population, instead of one,
the expected payoffs strictly yield a higher cooperation rate  even in the case
of the snowdrift game. Similar to the previous result, the difference between
the two approaches is always higher for the prisoner's dilemma class.

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=.33\textwidth]{static/rounds_two_by_two_games_cooperation.pdf}
  \includegraphics[width=.33\textwidth]{static/opponents_two_by_two_games_cooperation.pdf}
  \includegraphics[width=.33\textwidth]{static/rounds_opponents_two_by_two_games_cooperation.pdf}
  \caption{\textbf{Cooperation rates over various social dilemmas for different
  limited memory payoffs.} ({\bf A}) If
  players update based on their last two rounds with another member of the population.
  ({\bf B}) If players update based on their last round with two other members of the population.
  ({\bf C}) If players update based on their last two rounds with two other members of the population.
  We vary the temptation
  payoff \(T \in \{0.2, 0.6, 1, 1.4, 1.8, 2.2\}\)
  across the \(x\) axis, and \(S \in \{1.2, 0.8, 0.4, 0, -0.4, -0.8\}\)
  across the \(y\) axis. Unless explicitly varied, the parameters of the simulation
  are $N\!=\!100$, $\beta\!=\!10$, $\delta\!=\!0.99$, $t\!=\!5\times 10^7$..}
  \label{fig:cooperation_rates_two_by_two_part_two}
\end{figure}

\section{Conclusions}\label{section:conclusions}

Cooperation can be seen at odd, why is it that we choose to help others,
increasing their payoff, at the expense of decreasing one's own? In spite of all
the selfish genes’ animal and human communities seem to altruistically help each
other and cooperate, and evolutionary game theory has helped us shape our
understanding of the evolution of cooperation.

Previous evolutionary models often feature a curious inconsistency. When
modeling how individuals make decisions in each round, these models assume that
players only remember the last round. However, when modeling how individuals
update their strategies over time, individuals are assumed to have perfect
memory.

Here, we have explored how robust cooperation is as models deviate from the
perfect memory assumption. Initially we considered the donation game. We showed
that when the last round payoff is used instead of the expected payoffs,
cooperation can even evolve when individuals only use a minimum of information,
however, the evolving cooperation rates are typically lower. The resident
strategies were both less cooperative and less generous. This effect was only
intensified we increase the benefit and the strength of selection independently.
The results showed that as each parameter was being increased the difference
between the cooperation rates were widening. This indicates that cooperative
players benefit from being able to interact with everyone in the population.

We extended our approach and presented results not only based on different
classes of social games, but also based on different limited memory payoffs. The
analysis showed that specifically for the case of the prisoner's dilemma games
the contrast between the two approaches can be significantly large. The
prisoner's dilemma is one of the most applied types of social games when we
discuss the evolution of cooperation. Our results indicate that cooperation
struggles to evolve in the prisoner's dilemma when only a minimum of social
information is used at the updating stage. Interestingly our simulations also
showed that in same cases cooperation can benefit from limited memory. This was
specific for the snowdrift game.

\section{Acknowledgements}

This work was supported by the European Research Council Starting Grant 850529:
E-DIRECT.

\appendix

\section{Model Setup}\label{appendix:methods}

Consider a population of \(N\) individuals where \(N\) is even. At any point in
time there are at most two different strategies in present in the population.
More specifically, a mutant strategy played by \(k\) individuals and a resident
strategy played by \(N - k\) individuals. We assume a pairwise process in which
strategies spread because they are imitated more often. Each step of the
evolutionary process consists of two stages; a game stage and an update stage.

In the game stage, each individual is randomly matched with some other
individual in the population. Their interaction lasts for a number of turns
which is not fixed but depends on the continuation probability \(\delta\). At
each turn the individuals choose between cooperation (\(C\)) and defection
(\(D\)). Thus, there are four possible outcomes in each turn \(CC, CD, DC\) and
\(DD\). If both players cooperate they receive the reward payoff \(R\), whereas
if both players defect they receive the punishment payoff \(P\). If one
cooperates but the other defects, the defector receives the temptation to
defect, \(T\), whereas the cooperator receives the sucker's payoff, \(S\).
Let $\mathcal{U}=\{R,S,T,P\}$ denote the set of feasible payoffs in each round,
and let $\mathbf{u}\!=\!(R,S,T,P)$ be the corresponding payoff vector.
We present results for various values of $\mathcal{U}$ for all
the symmetric \(2 \times 2\) games.

A further assumption of our model is that individuals make use of reactive
strategies when they make decisions in each round. Reactive strategies are a set
of strategies that take into account only the previous action of the opponent.
A reactive strategy can be written explicitly as a vector,

\[s=(y, p, q)\]

where \(y\) is the probability that the strategy opens with a cooperation and
\(p, q\) are the probabilities that the strategy cooperates given that the
opponent cooperated and defected equivalently.

In the updating stage, two players are randomly drawn from the population, a
`learner' and a `role model'. The learner adopts the role model's strategy
based on the Fermi distribution function, %ToDo add reference

\begin{equation} \label{Eq:rho}
\rho(u_{L}, u_{RM}) = \frac{1}{1\!+\! \exp^{\!-\!\beta (u_{RM}\!-\!u_{L})}}.
\end{equation}

where $u_{L}\!\in\! \mathcal{U}$ is the learner's payoff, $u_{RM}\!\in\!
\mathcal{U}$ is the role model's payoff, and $\beta\!\ge\!0$ is the strength of
selection.

We iterate this basic evolutionary step until either the mutant strategy goes
extinct, or until it fixes in the population and becomes the new resident
strategy. After either outcome, we set $k$ to $1$ and we introduce a new mutant
strategy which is uniformly chosen from all reactive strategies at random.
Instead of simulating each step of the evolutionary process, we estimate the
probability that a newly introduced mutant fixes~\cite{nowak2004emergence}. This
is defined as the fixation probability of the mutant, and the standard form is
the following,

\begin{equation}\label{eq:appendix_fixation_probability}
\varphi = \frac{1}{1+\sum\limits_{i=1}^{N-1}\prod\limits_k^i \frac{\lambda^-_k}{\lambda^+_k}},
\end{equation}

where \(\lambda^-_k, \lambda^+_k\) are the probabilities that the number of
mutants decreases and increases respectively.

This process of mutation and fixation/extinction is iterated many times. The
evolutionary process is summarized by
Algorithm~\ref{algorithm:pairwise_comparison}.

\begin{algorithm}[!htbp]
  \SetAlgoLined
   $N \leftarrow$ population size\;
   $k \leftarrow 1$\;
   resident $\leftarrow (0, 0, 0)$\;
   \While{ $t <$ maximum number of steps}{
    mutant $\leftarrow$ random: $\{\emptyset \}\ \rightarrow R^3$\;
    fixation probability $\leftarrow \varphi $\;
    \If{$\varphi >$ random: $i \rightarrow [0, 1]$}{
      resident $\leftarrow$ mutant;
     }}
     \caption{Evolutionary process}\label{algorithm:pairwise_comparison}
\end{algorithm}

The aim of this work is to explore the effect of updating memory on the
cooperation rate of the evolved population. For this reason we consider two
different approaches when estimating the payoffs at the updating stage. The two
approaches we consider are those of (i) the expected and (ii) the limited memory
payoffs.

\subsection*{Expected Payoffs}

The expected payoffs are the conventional payoffs used in the updating
stage~\cite{imhof2010stochastic}. They are defined as the mean payoff of an
individual in a well-mixed population that engages in repeated games with all
other population members.

We first define the payoff of two reactive strategies at the game stage. Assume
two reactive strategies $s_1\!=\!(y_1, p_1, q_1$) and $s_2\!=\!(y_2,p_2,q_2)$.
It is not necessary to simulate the play move by move, instead the play between
the two strategies is defined a Markov matrix \(M\),

\begin{equation}\label{eq:transition_matrix}
  \input{static/matrix.tex}.
\end{equation}

whose stationary vector \(\mathbf{v}\), combined with the payoff \(u\), yields
the game stage outcome for each strategy,
\(\langle\mathbf{v}(s_1,s_2),\mathbf{u}\rangle\)~\cite{Hauert1997}.

In the updating stage the learner adopts the strategy of the role model based on
their updating payoffs. Given that there are only two different types in the
population at each time step we only need to define the expected payoff for a
resident (\(\pi_R\)) and for a mutant (\(\pi_M\)). Assume the resident strategy
\(s_R = (y_R, p_R, q_R)\) and the mutant strategy \(s_M = (y_M, p_M, q_M)\), the
expected payoffs are give by,

\begin{equation} \label{Eq:ExpPay}
  \begin{array}{lcrcr}
  \displaystyle \pi_R	&=	&\displaystyle \frac{N\!-\!k\!-\!1}{N-1}\cdot \langle\mathbf{v}(s_R,s_R),\mathbf{u}\rangle	&+	&\displaystyle\frac{k}{N-1}\cdot \langle\mathbf{v}(s_R,s_M),\mathbf{u}\rangle,\\[0.5cm]
  \displaystyle \pi_M	&=	&\displaystyle\frac{N-k}{N-1}\cdot \langle\mathbf{v}(s_M,s_R),\mathbf{u}\rangle&+	&\displaystyle\frac{k-1}{N-1}\cdot \langle\mathbf{v}(s_M,s_M),\mathbf{u}\rangle.\\
  \end{array}
\end{equation}

The number of mutant in the population increase if a learner resident adopts the
strategy of a mutant role model, and decreases if a mutant leaner adopts the
strategy of a resident. The probabilities that the number of mutants decreases
and increases, \(\lambda^-_k\) and \(\lambda^+_k\), are now explicitly defined
as,

\begin{align*} 
  \lambda^-_k &\!=\!\rho(\pi_R, \pi_M) \\
  \lambda^+_k &\!=\!\rho(\pi_M, \pi_R).
\end{align*}

\subsection*{Limited memory payoffs}

Initially, we discuss the case of the \textbf{last round updating payoff}. At the stage
game we define the payoff of a reactive strategy in the last round,
Proposition~\ref{proposition:last_round}.

\begin{Prop}\label{proposition:last_round}
    Consider a repeated game, with
    continuation probability $\delta$, between players with reactive strategies
    $s_1\!=\!(y_1, p_1, q_1$)  and $s_2\!=\!(y_2,p_2,q_2)$ respectively. Then the
    probability that the $s_1$ player receives the payoff $u\!\in\! \mathcal{U}$ in
    the very last round of the game is given by $v_{u}(s_1,s_2)$, as given by
    Equation~(\ref{Eq:LastRound}).

    \begin{equation} \label{Eq:LastRound}
      \setlength{\arraycolsep}{1pt}
      \begin{array}{rcl}

      v_{R}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)} \times R,\\[1cm]

      v_{S}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{y_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(q_1+r_1\big((1\!-\!\delta)y_2+\delta q_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)} \times S,\\[1cm]

      v_{T}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1y_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(q_2+r_2\big((1\!-\!\delta)y_1+\delta q_1\big)\Big)}
      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)} \times T,\\[1cm]

      v_{P}(s_1,s_2) &= &\displaystyle (1\!-\!\delta)\frac{\bar{y}_1\bar{y}_2}{1\!-\!\delta^2 r_1 r_2}+\delta \frac{\Big(\bar{q}_1+\bar{r}_1\big((1\!-\!\delta)y_2+\delta p_2\big)\Big) \Big(\bar{q}_2+\bar{r}_2\big((1\!-\!\delta)y_1+\delta p_1\big)\Big)}
      {\displaystyle(1\!-\!\delta r_1r_2)(1\!-\!\delta^2 r_1 r_2)} \times P.
      \end{array}
    \end{equation}

In these expressions, we have used the notation $r_i:=p_i\!-\!q_i$,
$\bar{y}_i\!=\!1\!-\!y_i$, $\bar{q}_i:=1\!-\!q_i$, and
$\bar{r}_i:=\bar{p}_i\!-\!\bar{q}_i=-r_i$ for $i\!\in\!\{1,2\}$.
\end{Prop}

\begin{proof}
Given a play between two reactive strategies with continuation probability
$\delta$. The outcome at turn \(t\) is given by,

\begin{equation}\label{eq:}
  (1 - \delta) \mathbf{v_0} \sum \delta^{t} M^{(t)},
\end{equation}

where $\mathbf{v_0}$ denotes the expected distribution of the four outcomes in
the very first round, and \(1- \delta\) the probability that the game ends.
It can be shown that,

\begin{align*}
  (1 - \delta) \mathbf{v_0} \sum \delta^{t} M^{(t)} & = (1 - \delta)(\mathbf{v_0} + \delta \mathbf{v_0} M + \delta^{2}\mathbf{v_0} M ^{2} + \dots )\\ 
   & = (1 - \delta)\mathbf{v_0} (1 + \delta M + \delta^{2}M ^{2} + \dots ) \text{ using standard formula for geometric series}\\ 
   & = (1 - \delta)\mathbf{v_0}(I_4 - \delta M)^{-1}
\end{align*}

where \((1 - \delta)\mathbf{v_0}(I_4 - \delta M)^{-1}\) is vector \(\in R^{4}\)
and it the probabilities for being in any of the outcomes \(CC, CD, DC, DD\) in
the last round. Combining this with the payoff vector \(u\) and some algebraic
manipulation we derive to the Equation~\ref{Eq:LastRound}.
\end{proof}

In the updating stage we select a mutant and resident to be either the role
model or the learner. Given that they can interact with only one other member of
the population, they can interact either with each other or either can interact
with another resident or with another mutant. Thus, in each updating stage there
are five possible combinations of pairs (Figure~\ref{fig:single_pairs}).

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=.65\textwidth]{static/last_rounds_matches.tex}
  \caption{\textbf{Possible pairings combination in the updating stage, given
  that individuals interact with only one other member in the population.} At
  each step of the evolutionary process we choose a role model and a learner to
  update the population. We consider the case where both the role model and the
  learner estimate their fitness after interacting with a single member of the
  population. There are five possible pairings at each step. They interact with
  other with a probability \(\frac{1}{N - 1}\), and thus they do not interact
  with other with a probability \(1 - \frac{1}{N - 1}\). In
  the latter case, each of them can interact with either a mutant or a resident.
  Both of them interact with a mutant with a probability
  $\frac{(k-1)(k-2)}{(N-2)(N-3)}$ and both interact with a resident with a
  probability $\frac{(N-k-1)(N-k-2)}{(N-2)(N-3)}$. The last two possible pairings
  are that either of them interacts with a resident whilst the other interacts
  with a mutant, and this happens with a probability $\frac{(N-k-1)(k-1)}{(N-2)(N-3)}$.}
  \label{fig:single_pairs}
\end{figure}

Given the last round payoff and possible pair combinations for a single interaction,
we define the probability that the respective last round payoffs of two players
\(s_1, s_2\) are given by $u_1$ and $u_2$ as,

\begin{equation}\label{eq:Chi}
\setlength{\arraycolsep}{1pt}
\begin{array}{llrl}
x(u_1,u_2)	 &=&\displaystyle \frac{1}{N\!-\!1}\cdot  &v_{u_1}(s_1,s_2)\cdot 1_{(u_1,u_2)\in \mathcal{U}^2_F}\\[0.5cm]
&+	
&\displaystyle \left(1\!-\!\frac{1}{N\!-\!1}\right)  
&\left[ \frac{k\!-\!1}{N\!-\!2}\frac{k\!-\!2}{N\!-\!3} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_2) + 
 \frac{k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!1}{N\!-\!3} v_{u_1}(s_1,s_2) v_{u_2}(s_2,s_1)\right.\\[0.5cm]
&&&\left. +\frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{k\!-\!1}{N\!-\!3} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_2) + 
 \frac{N\!-\!k\!-\!1}{N\!-\!2}\frac{N\!-\!k\!-\!2}{N\!-\!3} v_{u_1}(s_1,s_1) v_{u_2}(s_2,s_1)\right].
\end{array}
\end{equation}

The first term on the right side corresponds to the case that the learner and
the role model happened to be matched during the game stage, which happens with
probability $\frac{1}{(N\!-\!1)}$. In that case, we note that only those payoff pairs
can occur that are feasible in a direct interaction, $\splitatcommas{(u_1,u_2)\in
\mathcal{U}^2_F:=\big\{ (R,R), (S,T), (T,S), (P,P) \big\}}$, as represented by
the respective indicator function. Otherwise, if the learner and the role model
did not interact directly, we need to distinguish four different cases,
depending on whether the learner was matched with a resident or a mutant, and
depending on whether the role model was matched with a resident or a mutant.

Given that $N\!-\!k$ players use the resident strategy $s_{R}\!=\!(y_{R},p_{R},q_{R})$
and that the remaining $k$ players use the mutant strategy
$s_{M}\!=\!(y_{M},p_{M},q_{M})$, the probability that the number of mutants increases by
one in one step of the evolutionary process can be written as

\begin{align}
\lambda^+_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M})\cdot \rho(u_{R},u_{M}), \\
\lambda^-_k=\frac{N\!-\!k}{N}\cdot \frac{k}{N}\cdot \sum_{u_{R},u_{M}\in\mathcal{U}} x(u_{R},u_{M})\cdot \rho(u_{M},u_{R}).
\end{align}

In this expression, $\frac{(N\!-\!k)}{N}$ is the probability that the randomly chosen
learner is a resident, and $\frac{k}{N}$ is the probability that the role model is a
mutant. The sum corresponds to the total probability that the learner adopts the
role model's strategy over all possible payoffs $u_R$ and $u_M$ that the two
player may have received in their respective last rounds. We use $x(u_R,u_M)$ to
denote the probability that the randomly chosen resident obtained a payoff of
$u_R$ in the last round of his respective game, and that the mutant obtained a
payoff of $u_M$.

We extend our framework to consider the case where players update their
strategies based on the outcome of \textbf{the last two turns and based on their
interaction with two other members of the population}. At the stage game we
define the payoff of a reactive strategy in the last two rounds,
Proposition~\ref{proposition:last_two_rounds}.

\begin{Prop}\label{proposition:last_two_rounds}
  Consider a repeated game, with continuation probability $\delta$, between
  players with reactive strategies $s_1\!=\!(y_1, p_1, q_1$)  and
  $s_2\!=\!(y_2,p_2,q_2)$ respectively. Let $\mathcal{\tilde{U}} =
  \splitatcommas{\{RR, RS, RT, RP, SR, SS, ST, SP, TR, TS, TT, TP, PR, PS, PT,
  PP\}}$ denote the set of feasible payoffs in the last two rounds, and let
  \(\tilde{\mathbf{u}}\) be the corresponding payoff vector. Then the
  probability that the $s_1$ player receives the payoff $u\!\in\!
  \mathcal{\tilde{U}}$ in the very last two rounds of the game is given by,

  \begin{equation}
  \langle\mathbf{\tilde{v}}(s_1,s_2),\mathbf{\tilde{u}}\rangle, \text{ where } \mathbf{\tilde{v}} \in R^{16} \text{ is given by },
  \end{equation}
  \begin{equation}
    \mathbf{\tilde{v}}(s_1,s_2) = (1 - \delta) m_{a_1, a_2} \delta^2 \left[\mathbf{v_0}(I_4 - \delta M)^{-1}\right]_{a_1, a_2}, \quad  m_{a_1, a_2} \in M \ \forall \ a_1, a_2 \in \{1, 2, 3, 4\}.
  \end{equation}
\end{Prop}

In the updating stage we select a mutant and resident to be either the role
model or the learner. Given that they can interact with two other members of
the population there are a total of twenty four possible combinations of pairs
(Figure~\ref{fig:pissible_two_pairs}).

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=.65\textwidth]{static/tree.tex}
  \caption{\textbf{Possible pairings combination in the updating stage, given
  that individuals interact with two other members in the population.}}
  \label{fig:pissible_two_pairs}
\end{figure}

Given the last two rounds payoff and possible pair combinations with two
members, we can define the probability \(x\) that the respective last round
payoffs of two players \(s_1, s_2\) are given by $u_1$ and $u_2$ similarly to
Eq.~(\ref{eq:Chi}). We follow the same approach to define the rest of the
updating payoffs. These are the updating payoffs of the last round with two
member, and the the last two rounds payoffs with one member.

Simulating the evolutionary process for more interactions and rounds quickly
becomes computationally intractable. Our methodology could be extended to
include \(n\) turns and \(m\) interactions. However, for the purpose of this
work we explore the cases only up to two turns and two interactions.

\section{Evolutionary dynamics under limited memory payoffs for various social dilemmas}\label{appendix:further_simulation_results}

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\textwidth]{static/rounds_two_by_two_games.pdf}
  \caption{{\bf Evolutionary dynamics under last two rounds payoffs for various social dilemmas.} 
  We have run several simulations of the evolutionary process described in
  section~\ref{section:model} for $t\!=\!10^7$ time steps. The graphs show how
  often the resident population chooses each combination ($p,q$) of conditional
  cooperation probabilities in the subsequent rounds. We vary the temptation
  payoff \(T \in \{-1, -0.6, -0.2,  0.2, 0.6, 1, 1.4, 1.8, 2.2, 2.6, 3\}\)
  across the \(x\) axis, and  \(S \in \{2, 1.6, 1.2, 0.8, 0.4, 0, -0.4, -0.8,
  -1.2, -1.6, -2\}\) across the \(y\) axis. Parameters: $N\!=\!100$,
  $\beta\!=\!10$, $\delta\!=\!0.999$.}
  \label{fig:last_two_rounds_two_by_two}
\end{figure}

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\textwidth]{static/opponents_two_by_two_games.pdf}
  \caption{{\bf Evolutionary dynamics under last round payoffs with two members of the population for various social dilemmas.} 
  We have run several simulations of the evolutionary process described in
  section~\ref{section:model} for $t\!=\!10^7$ time steps. The graphs show how
  often the resident population chooses each combination ($p,q$) of conditional
  cooperation probabilities in the subsequent rounds. We vary the temptation
  payoff \(T \in \{-1, -0.6, -0.2,  0.2, 0.6, 1, 1.4, 1.8, 2.2, 2.6, 3\}\)
  across the \(x\) axis, and  \(S \in \{2, 1.6, 1.2, 0.8, 0.4, 0, -0.4, -0.8,
  -1.2, -1.6, -2\}\) across the \(y\) axis. Parameters: $N\!=\!100$,
  $\beta\!=\!10$, $\delta\!=\!0.999$.}
  \label{fig:last_two_opponents_two_by_two}
\end{figure}

\begin{figure}[!htbp]
  \centering
  \includegraphics[width=\textwidth]{static/merged_plot_rounds_opponents_two.pdf}
  \caption{{\bf Evolutionary dynamics under last two rounds payoffs with two members of the population for various social dilemmas.} 
  We have run several simulations of the evolutionary process described in
  section~\ref{section:model} for $t\!=\!10^7$ time steps. The graphs show how
  often the resident population chooses each combination ($p,q$) of conditional
  cooperation probabilities in the subsequent rounds. We vary the temptation
  payoff \(T \in \{-1, -0.6, -0.2,  0.2, 0.6, 1, 1.4, 1.8, 2.2, 2.6, 3\}\)
  across the \(x\) axis, and  \(S \in \{2, 1.6, 1.2, 0.8, 0.4, 0, -0.4, -0.8,
  -1.2, -1.6, -2\}\) across the \(y\) axis. Parameters: $N\!=\!100$,
  $\beta\!=\!10$, $\delta\!=\!0.999$.}
  \label{fig:last_two_rounds_two_opponents}
\end{figure}

\bibliographystyle{unsrt}
\bibliography{bibliography.bib}

\end{document}