### 2021-06-04 static/expected_stochastic.tex
Draft paper (#5)
* compile latex

* re write simulations.py to check that the simulated utility is equal to the theoretical

* initial analysis of data

* fixes bug in matlab code

The current code wrote the cooperation rate for the mutant at
each time step instead of the rate of the resident

* run analysis

* check new data output

* use csvwrite instead of writematrix

* tweaks analysis code to fit the new data

* add the functionality to consider more than just the last turn

* scoring turns as a class argument

* test function further

* run black

* some style fixes

* matlab scripts for

\documentclass{standalone}
\usepackage{tikz}

\begin{document}
\begin{tikzpicture}
    \draw[->,ultra thick] (0, 0)--(6,0) node[below left]{$n$};
    \draw[->,ultra thick] (0, 0)--(0,6) node[below left]{$m$};

    \node[fill, circle, label=above:expected, inner sep=2pt] at (5.7, 5.7) {};
    \node[fill, circle, label=above:stochastic, inner sep=2pt] (stochastic) at (1, 1) {};
    \node[label=left:1] (y) at (0, 1) {};
    \node[label=below:1] (x) at (1, 0) {};

    \draw[dotted] (y)--(1, 1) {};
    \draw[dotted] (x)--(1, 1) {};
\end{tikzpicture}
\end{document}