### 2022-08-26 paper/population.tex

\documentclass{standalone}

\begin{document}
\tikzset{man/.pic={
    \node[circle,fill,minimum size=5mm] (head) {};
    \node[rounded corners=2pt,minimum height=.8cm,minimum width=0.4cm,fill,below = 1pt of head] (body) {};
    \draw[line width=1mm,round cap-round cap] ([shift={(2pt,-1pt)}]body.north east) --++(-90:6mm);
    \draw[line width=1mm,round cap-round cap] ([shift={(-2pt,-1pt)}]body.north west)--++(-90:6mm);
    }}
\begin{tikzpicture}

    \node at (-.7, 2) {A)};
    \pic[Apricot] at (0, 2) (myman) {man};
    \node at (0, 1.4) {A};
    \pic[BurntOrange] at (1, 2) (myman) {man};
    \node at (1, 1.4) {B};
    \pic[OliveGreen] at (2, 2) (myman) {man};
    \node at (2, 1.4) {C};

    \pic[Orchid] at (0, 0) (myman) {man};
    \node at (0, -.6) {E};
    \pic[RoyalBlue] at (1, 0) (myman) {man};
    \node at (1, -.6) {F};
    \pic[Maroon] at (2, 0) (myman) {man};
    \node at (2, -.6) {D};

    \node at (1, -2) {$\dots$};

\end{tikzpicture}
\end{document}