### 2023-02-21 paper/static/mutation.tex

\documentclass{standalone}

\begin{document}
\tikzset{man/.pic={
    \node[circle,fill,minimum size=5mm] (head) {};
    \node[rounded corners=2pt,minimum height=.8cm,minimum width=0.4cm,fill,below = 1pt of head] (body) {};
    \draw[line width=1mm,round cap-round cap] ([shift={(2pt,-1pt)}]body.north east) --++(-90:6mm);
    \draw[line width=1mm,round cap-round cap] ([shift={(-2pt,-1pt)}]body.north west)--++(-90:6mm);
    }}
\begin{tikzpicture}

    %%%%% mutation %%%%%
    \pic[Apricot] at (0, 0) (myman) {man};
    \node at (0, -.6) {A};
    \pic[Goldenrod] at (1.5, 0) (myman) {man};
    \node at (1.5, -.6) {A};

    \coordinate (c1) at (.5, -.5);
    \coordinate (c2) at (1, -.5);
    \draw (c1) edge[out=15, in=165, ->, thick] node [above] {$\mu$} (c2);

\end{tikzpicture}
\end{document}