### 2023-11-02 paper/static/person.tex

\documentclass[tikz]{standalone}

\definecolor{solarizedBlue}{HTML}{268BD2}
\definecolor{bostonuniversityred}{rgb}{0.8, 0.0, 0.0}
\definecolor{ao}{rgb}{0.0, 0.5, 0.0}

\usetikzlibrary{positioning, fit, calc,matrix, arrows,scopes}

\begin{document}
\tikzset{man/.pic={
    \node[circle,fill,minimum size=10mm] (head) {};
    \node[rounded corners=4pt,minimum height=1.1cm,minimum width=0.8cm,fill,below = 2pt of head] (body) {};
    \draw[line width=2mm,round cap-round cap] ([shift={(4pt,-2pt)}]body.north east) --++(-90:6mm);
    \draw[line width=2mm,round cap-round cap] ([shift={(-4pt,-2pt)}]body.north west)--++(-90:6mm);
    }}
\begin{tikzpicture}

\pic[ao] at (0, 0) (myman) {man};

\end{tikzpicture}
\end{document}